\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2017-01-20 mSystems/rev2/mSys_MAIN_rev2.tex
mSys: Kick off for REV2 (DRAFT rel alpha)
Changes to be committed:
	new file:   ../documentation/DetailedEditorAnswer-R1.pdf
	renamed:    ../rev/Response to Reviewer Comments.docx -> ../rev1/Response to Reviewer Comments.docx
	renamed:    ../rev/figs/Fig1.eps -> ../rev1/figs/Fig1.eps
	renamed:    ../rev/figs/Fig2.eps -> ../rev1/figs/Fig2.eps
	renamed:    ../rev/figs/Fig2/Fig2.pdf -> ../rev1/figs/Fig2/Fig2.pdf
	renamed:    ../rev/figs/Fig2/Fig2.svg -> ../rev1/figs/Fig2/Fig2.svg
	renamed:    ../rev/figs/Fig2/README -> ../rev1/figs/Fig2/README
	renamed:    ../rev/figs/Fig2/ellipse.m -> ../rev1/fi

% watermarks (time consuming!)
%\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
%\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
%\emph{mSystems} \LaTeX\ manuscript \releasecode}

@@

%===================================== SECTIONS =======================================
%%% Introduction
\section*{Introduction}
\input{mSys_intro_rev.tex}
%\clearpage
%%% Result
\section*{Results}
\input{mSys_results_rev.tex}
%\clearpage
%%% Discussion
\section*{Discussion}
\input{mSys_discussion_rev.tex}
%\clearpage    
%%% Material & Methods
\section*{Materials and Methods}
\input{mSys_MandM_rev.tex}   
%\clearpage

@@

%% Results
\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
% -> main studies
\bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011. Moving pictures of the human microbiome. Genome Biol {\bf 12}:R50.
\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. Proc Natl Acad Sci {\bf108}:4554–61.
\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013. The long-term stability of the human gut microbiota. Science {\bf 341}:1237439.
\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. Science {\bf 339}:548–54.
\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014. Diet rapidly and reproducibly alters the human gut microbiome. Nature {\bf 505}:559–63.
\bibitem{hostlife} {\bf David LA, Materna AC, Friedman J, Campos-Baptista MI, Blackburn MC, Perrotta A, Erdman SE, Alm EJ.} 2014. Host lifestyle affects human microbiota on daily timescales. Genome Biol {\bf 15}:R89.
% <- END main studies
\bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.
\bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83}:066115.
\bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.

>>>>

% watermarks (time consuming!)
\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript \releasecode}

@@

%===================================== SECTIONS =======================================
%%% Introduction
\section*{Introduction}
\input{mSys_intro_rev2.tex}
%\clearpage
%%% Result
\section*{Results}
\input{mSys_results_rev2.tex}
%\clearpage
%%% Discussion
\section*{Discussion}
\input{mSys_discussion_rev2.tex}
%\clearpage    
%%% Material & Methods
\section*{Materials and Methods}
\input{mSys_MandM_rev2.tex}   
%\clearpage

@@

%% Results
\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
\bibitem{medicare} {\bf Hidalgo CA, Blumm N, Barabási AL, Christakis NA.} 2009. A Dynamic Network Approach for the Study of Human Phenotypes. PLoS Comput Biol {5}.
% -> main studies
\bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011. Moving pictures of the human microbiome. Genome Biol {\bf 12}:R50.
\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. Proc Natl Acad Sci {\bf108}:4554–61.
\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013. The long-term stability of the human gut microbiota. Science {\bf 341}:1237439.
\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. Science {\bf 339}:548–54.
\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014. Diet rapidly and reproducibly alters the human gut microbiome. Nature {\bf 505}:559–63.
\bibitem{hostlife} {\bf David LA, Materna AC, Friedman J, Campos-Baptista MI, Blackburn MC, Perrotta A, Erdman SE, Alm EJ.} 2014. Host lifestyle affects human microbiota on daily timescales. Genome Biol {\bf 15}:R89.
% <- END main studies
\bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.
\bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83}:066115.
\bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.

### 2017-01-20 mSystems/rev2/mSys_intro_rev2.tex
causality and medicare factoid points fixed in the text

This paper presents the imprints of health status (healthy or disease) in the macroscopic properties of microbiota, by studying its temporal variability. We analyzed more than 40000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, trips, obese status, antibiotic therapy and healthy individuals. Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}. We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this affected the healthy status of the subjects.

>>>>

This paper presents the imprints of health status (healthy or disease) in the macroscopic properties of microbiota, by studying its temporal variability. We analyzed more than 40000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, trips, obese status, antibiotic therapy and healthy individuals. Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}. We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this is related to the healthy status of the subjects.
